EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-04-14 05_examples.tex
Update. Start drafting Examples:AHM, getting data Fortan and Python (thanks Lorenzo Crippa)
co-authored-by: Lorenzo Crippa <lcrippa@gmail.com>

\subsection{Bethe lattice DMFT (Fortran API)}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$.

We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
defined on a Bethe lattice with density of states
$\rho(x)=\frac{1}{2D}\sqrt{D^2-x^2}$, where $D=2t$ is the
half-bandwidth.

@@

\subsection{Attractive Hubbard model (Python API)}
\subsection{Multi-orbital Hubbard (Triqs)}

>>>>

\subsection{Bethe lattice DMFT (Fortran API)}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$. 
We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
defined on a Bethe lattice with density of states
$\rho(x)=\frac{1}{2D}\sqrt{D^2-x^2}$, where $D=2t$ is the
half-bandwidth~\cite{Georges1996}.
In DMFT this model is mapped
onto a quantum impurity problem with an effective electronic bath
which needs to be determined self-consistently.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   !> Number of discretization points for the Bethe DOS
   integer,parameter                           :: Le=5000
   !> Bethe half-bandwidth = energy unit
   real(8),parameter                           :: D=1d0
   !>Bath:
   real(8),allocatable                         :: Bath(:)
   !> local dynamical functions, rank-5 [Nspin,Nspin,Norb,Norb,L]
   complex(8),allocatable,dimension(:,:,:,:,:) :: Weiss,Smats,Sreal,Gmats,Greal,Weiss_
   ... (omissis)
   !> READ THE input using EDIpack procedure:
   call ed_read_input(trim(finput))
   !
   !> Setup the Bethe DOS $\rho$ (linear dispersion)
   Ene = linspace(-D,D,Le,mesh=de)
   DOS = dens_bethe(Ene,D)
\end{lstlisting}

The bath is described
by the (unknown) function $\GG^{-1}_0$, i.e. the Weiss field. Within \NAME
exact diagonalization approach, where the bath is approximated with a
finite number of levels, the function $\GG^{-1}_0$ is used to
determine bath parameters $\vec{x}=\{V,h\}$ using the optimization method discussed in
\secu{sSecFit}.
The starting point of any calculation is determined by a suitable
guess of the Weiss field, or equivalently of the bath parameters.
In \NAME this operations is accomplished by the function {\tt
  ed\_init\_solver}.

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
Nb=ed_get_bath_dimension()
allocate(bath(Nb))
call ed_init_solver(bath)
\end{lstlisting}

The iterative algorithm proceeds as follows:
\begin{enumerate}
\item Call the exact diagonalization {\bf impurity solver} {\tt
    ed\_solve} whose only input is the set of parameters $\vec{x}$. All other
  options are controlled by input file specifications ({\tt
    ed\_read\_input}).
\item Retrieve at least the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using {\tt ed\_get\_sigma}.
\item Evaluate the local interacting Green's function
  $G_{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e$ with
  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$.
  \item Update the Weiss field through the {\bf self-consistency}
    relation: $\GG^{-1}_0(i\omega) = G_{loc}(i\omega) + \Sigma(i\omega)$. 
  \item Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field using, for instance, {\tt ed\_chi2\_fitgf}. Restart at 1.
\end{enumerate}
In agreement with Reverse Communication Strategy setup, the step 3-4
are left to the user. In addition step 5 is kept separated to
guarantee enough freedom in the bath optimization, though we remark
this is a crucial step which can potentially compromise the whole
calculation.
An example of implementation is given in the following listings.

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
     iloop=iloop+1     
     !> Solve the effective impurity problem
     call ed_solve(bath)     
     !> Retrieve impurity Self-energy on Matsubara axis
     call ed_get_sigma(Smats,'m')
     !
     !> Build a local Green's function:
     wfreq = pi/beta*(2*arange(1,Lmats)-1)   !automatic Fortran allocation
     do i=1,Lmats ! One can do better than this of course 
        zeta= xi*wfreq(i)+xmu - Smats(1,1,1,1,i)
        Gmats(1,1,1,1,i) = sum(DOS(:)/( zeta-Ene(:) ))*de  
     enddo
     !> Update Weiss field: Self-consistency
     Weiss(1,1,1,1,:) = one/(one/Gmats(1,1,1,1,:) + Smats(1,1,1,1,:))
     !> Close the self-consistency fitting the new bath:
     call ed_chi2_fitgf(Weiss,bath,ispin=1)     
     !>Check convergence
     converged =( sum(abs(Weiss(1,1,1,1,:)-Weiss_(1,1,1,1,:)))/sum(abs(Weiss(1,1,1,1,:))) )<dmft_error
     Weiss_=Weiss     
   enddo
\end{lstlisting}

\subsubsection{Results}
In the following we present some results obtained by executing this
simple program varying the interaction strenght :f:var:`uloc`.
Differently from the previous case of a quantum impurity embedded in a
given bath describing the progressive formation of a
strongly renormalized Fermi liquid state, here the DMFT
self-consistency allows to describe the transition from a correlated
metal to a Mott insulating state.

To illustrate this point, in panel **A** we report the evolution of
the spectral function $-{\rm Im}G(\omega)/\pi$ as a function of
$U$. Despite the *spiky* nature of the spectrum, due to the
finite size (i.e. number of poles) of the discretized effective bath,
one can clearly distinguish the renormalization of the central
quasi-particle peak at low-energy and the concomitant formation of
rather incoherent  high-energy features which will develop into
Hubbard bands for $U>U_c$, with $U_c\simeq 2.8D$.

@@

In the panels (B) and (C) we further discuss the metal-insulator
transition by showing the evolution of the self-energy functions.
In panel (B) we report the self-energy ${\rm Im}\Sigma(i\omega)$
on the Matsubara axis in the low energy regime. Increasing $U$
we observe the progressive growth of this function until it takes a
diverging behavior crossing the critical interaction strenght. We
recall that this behavior can be observed on the Matsubara axis
because of the particle-hole symmetry of the problem.

Using the relation:

$$
   \frac{\Im\Sigma(i\omega_n)}{\omega_n}_{|_{\omega_n\rightarrow 0}}=
   \frac{1}{\pi}\int_{\mathbb R}d\epsilon \frac{\Re\Sigma(\epsilon)}{\epsilon^2}=
   \frac{\partial\Re\Sigma}{\partial\omega}_{|_{\omega\rightarrow 0}}.
$$
we can extract the quasi-particle renormalization constant :math:`Z`
from the linear behavior of ${\rm Im}\Sigma(i\omega)$ near
$\omega=0$ in the metallic regime. The results are highlighted
in the figure and the values of $Z$ are reported in the legend.

Using the right hand side of the previous relation, we show in
panel (C) the behavior of  ${\rm Re}\Sigma(\omega)$
on the real axis around the Fermi level. Again, increasing $U$
we observe the slope of the linear behavior to increase until the
critical point is crossed and an insulating state is reached. On the
real-axis this is signaled by the divergence of the imaginary part of
the self-energy   ${\rm Im}\Sigma(\omega) \rightarrow -\infty$
near the chemical potential, which here is set to zero by
particle-hole symmetry. The corresponding real part shows a
discontinuity visibile in the panel (C).

\subsection{Attractive Hubbard model (Python API)}
The second application has the twofold aim to demonstrate the
superconductivity support of the \NAME library as well as to
illustrate the Python API through a working example.

\subsection{Multi-orbital Hubbard (Triqs)}

\subsection{Some model (W2Dynamics)}

\subsection{Interacting Bernevig-Hughes-Zhang (Fortran)}
